---
title: "CEO Compensation Appraisal Tool Results - Version 01"
output: pdf_document
params:
  org: NA
  search: NA
  hard: NA
  testorg: NA
---



```{r HEOM, echo=FALSE, warning=F, message=F}
library(knitr)

#Set wd to be repo folder 
setwd("..")

### See pretend-shiny vignette for detailed information

#get parameters 
org <- params$org
search <- params$search
hard <- params$hard
test.org <- params$testorg

#get filtered data
dat.filtered <- dat_filtering_hard(search, hard, test.org$EIN)

#get results 
all.results <- HEOM_with_weights(org, search, dat.filtered)
results <- all.results[[1]]
distance.matrix <- all.results[[2]]



#getting weighted average 
d_i <- results$dist
sum_w <- sum(1 - d_i)
w_i <- (1-d_i) / sum_w
#sum(w_i) = 1

weighted.average <- sum(w_i * results$CEOCompensation)

#getting the pay range
dat.model <- results

#make everything factors
dat.model$FormYr <- as.factor(dat.model$FormYr )
dat.model$MajorGroup <- as.factor(dat.model$MajorGroup)
dat.model$UNIV <- as.factor(dat.model$UNIV )
dat.model$HOSP <- as.factor(dat.model$HOSP )
dat.model$State <- as.factor(dat.model$State)
dat.model$Loc <- as.factor(dat.model$LocationType )



#only use complete data
dat.model <- na.omit(dat.model)

form <- "CEOCompensation ~ FormYr + log(TotalAssests+1, 10) + log(GrossReceipts + 1, 10) + log(TotalExpense+1, 10) + log(TotalEmployee+1, 10)"

#if there is more than one major group, add it
if(length(unique(search$MajorGroup)) > 1){
  form <- paste(form, " + MajorGroup")
}

#if there is more than one state, add it
if(length(unique(search$State)) > 1){
  form <- paste(form, " + State")
}

#if we are considering Universities and non-Universities, include UNIV
if(search$UNIV == 2){
  form <- paste(form, " + UNIV")
}

#if we are considering hospitals and non-hospitals, include HOSP
if(search$HOSP == 2){
  form <- paste(form, " + HOSP")
}


#if we are considering both Rural and Metro
if(length(unique(search$Loc)) > 1){
  form <- paste(form, " + Loc")
}

#if both male and female in the results set
if(all( c("F", "M") %in% unique(results$Gender))){
  #filter out unknown gender
  dat.model <- dat.model%>%
    filter(Gender != "U")
  #update form
  form <- paste(form, " + Gender")
}



#add location and gender

form <- stats::formula(form)
mod <- lm(form , data = dat.model )

#get residuals as a percentage of their original value
resids <- mod$residuals
per <- resids / dat.model$CEOCompensation / 100
quants <- quantile(per, c(0.10, 0.90)) #get rid of the bonus years

suggested.range <- weighted.average + weighted.average * quants

pay.min <- dollarize(suggested.range[1])
pay.max <- dollarize(suggested.range[2])



# other characteristics we need
n <- dim(dat.filtered)[1]



tab.major.group <- table(dat.filtered$MajorGroup)
tab.ntee <- table(dat.filtered$NTEE)

```

## Suggested Pay

Your suggested CEO total pay range is **`r pay.min` - `r pay.max`**. 

Your weighted average pay is **`r weighted.average`**.

Your actual pay was **`r test.org$CEOCompensation`**.

## Your Inputs

Here we list the inputs that you provided. 

```{r OrgTable, echo=FALSE, message=FALSE, warning=FALSE}

get.major.group <- function(mg){
  ret <- case_when(mg == 1 ~ "Arts, Culture, and Humanities",
                   mg == 2 ~ "Education",
                   mg == 3 ~ "Environment and Animals",
                   mg == 4 ~ "Health",
                   mg == 5 ~ "Human Services",
                   mg == 6 ~ "International, Foreign Affairs",
                   mg == 7 ~ "Public, Societal Benefit",
                   mg == 8 ~ "Religion Related",
                   mg == 9 ~ "Mutual/Membership Benefit",
                   mg == 10 ~ "Unknown/Unclassified",
                   mg == 11 ~ "Hosptial",
                   mg == 12 ~ "University")
  return(ret)
  
}

get.ntee <- function(ntee){
  
  ret <- case_when(ntee == "D" ~ "Animals" ,
                   ntee == "A" ~ "Arts, Culture, and Humanities", 
                   ntee == "R" ~ "Civil Rights, Social Action & Advocacy", 
                   ntee == "S" ~ "Community Improvement & Capacity Building", 
                   ntee == "I" ~ "Crime & Legal-Related", 
                   ntee == "B" ~ "Education", 
                   ntee == "C" ~ "Environment",
                   ntee == "J" ~ "Employment", 
                   ntee == "K" ~ "Food, Agriculture & Nutrition", 
                   ntee == "E" ~ "Health Care",
                   ntee == "L" ~ "Housing & Shelter", 
                   ntee == "P" ~ "Human Services", 
                   ntee == "Q" ~ "International, Foreign Affairs & National Security",
                   ntee == "H" ~ "Medical Research", 
                   ntee == "F" ~ "Mental Health & Crisis Intervention", 
                   ntee == "Y" ~ "Mutual & Membership Benefit", 
                   ntee == "T" ~ "Philanthropy, Voluntarism & Grantmaking Foundations", 
                   ntee == "M" ~ "Public Safety, Disaster Preparedness & Relief", 
                   ntee == "W" ~ "Public & Societal Benefit", 
                   ntee == "G" ~ "Voluntary Health Associations & Medical Disciplines", 
                   ntee == "N" ~ "Recreation & Sports", 
                   ntee == "X" ~ "Religion Related", 
                   ntee == "U" ~ "Science & Technology", 
                   ntee == "V" ~ "Social Science", 
                   ntee == "O" ~ "Youth Development", 
                   ntee == "Z" ~ "Unknown, Unclassified")
  
  return(ret)
  
}


get.ntee.cc <- function(cc){
  
   ret <- case_when(cc == "01" ~ "Alliance/Advocacy Organization", 
                    cc == "02" ~ "Management and Technical Assistance", 
                    cc == "03" ~ "Professional Society/Association", 
                    cc == "05" ~ "Research Institute and/or Public Policy Analysis", 
                    cc == "11" ~ "Monetary Support - Single Organization", 
                    cc == "12" ~ "Monetary Support - Multiple Organizations", 
                    cc == "19" ~ "Nonmonetary Support Not Elsewhere Classified (N.E.C.)",
                    cc == "00" ~ "Regular nonprofit", 
                    is.na(cc) ~ "Regular nonprofit")
   return(ret)
}


org.table <- data.frame(Attributes = c("Name",
                                       "State", 
                                       "City Type", 
                                       "Broad Category", 
                                       "Major Group", 
                                       "Specality Orginization?", 
                                       "Hospital?", 
                                       "University?", 
                                       "Total Expenses", 
                                       "Total Employees"), 
                        Input = c(test.org$Name, 
                                  org$State,
                                  org$Loc,
                                  get.major.group(org$MajorGroup), 
                                  get.ntee(org$NTEE),
                                  get.ntee.cc(org$NTEE.CC), 
                                  ifelse(org$HOSP == TRUE, "Yes", "No"), 
                                  ifelse(org$UNIV == TRUE, "Yes", "No"),
                                  dollarize(org$TotalExpense),
                                  org$TotalEmployee))


org.table %>%
  knitr::kable(col.names = c("Attribute", "Input"),
               caption = "Your Orginization Inputs",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")

```



```{r SearchTable, echo=FALSE, message=FALSE, warning=FALSE}
search.table <- data.frame(
  Attributes = c("State", 
                 "City Type", 
                 "Broad Category", 
                 "Major Group", 
                 "Specality Orginization?", 
                 "Hospital?", 
                 "University?", 
                 "Total Expenses", 
                 "Total Employees"), 
  Input = c(paste(search$State, collapse = "; "),
            paste(search$Loc, collapse = ";"),
            paste(unlist(lapply(search$MajorGroup, get.major.group )), collapse = "; "), 
            paste(unlist(lapply(search$NTEE, get.ntee )), collapse = "; "),
            paste(unlist(lapply(search$NTEE.CC, get.ntee.cc )), collapse = "; "),
            ifelse(search$HOSP == TRUE, "Yes", "No"), 
            ifelse(search$UNIV == TRUE, "Yes", "No"),
            paste(dollarize(search$TotalExpense)[1], "to", dollarize(search$TotalExpense[2])),
            paste(search$TotalEmployee[1], "to", search$TotalEmployee[2])), 
  Hard = c(ifelse(hard$State == TRUE, "Hard", "Soft"), 
           ifelse(hard$Loc == TRUE, "Hard", "Soft"),
           ifelse(hard$MajorGroup == TRUE, "Hard", "Soft"),
           ifelse(hard$NTEE == TRUE, "Hard", "Soft"),
           ifelse(hard$NTEE.CC == TRUE, "Hard", "Soft"),
           "N/A", 
           "N/A", 
           ifelse(hard$TotalExpense == TRUE, "Hard", "Soft"),
           ifelse(hard$TotalEmployee == TRUE, "Hard", "Soft")))
search.table %>%
  knitr::kable(col.names = c("Attribute", "Search Critera", "Hard or Soft Filtering"),
               caption = "Your Search Criteria",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = c("hold_position","scale_down"))


```



## Characteristics of Your Comparison Set

There were **`r n`** other non-profits in your comparison set. 

```{r MajorGroup, echo=FALSE, warning=F, message=F}
tab.temp.mg <- tab.major.group %>% as.data.frame()

colnames(tab.temp.mg) <- c("MajorGroup", "Freq")

tab.temp.mg %>%
  as.data.frame() %>%
  dplyr::mutate(MajorGroup = dplyr::recode(MajorGroup,
                               `1` = "Arts, Culture, and Humanities",
                                `2` =  "Education",
                                `3` = "Environment and Animals" ,
                                `4` = "Health",
                                `5` = "Human Services",
                                `6` = "International, Foreign Affairs",
                                `7` = "Public, Societal Benefit",
                                `8` = "Religion Related",
                                `9` = "Mutual/Membership Benefit",
                                `10` = "Unknown/Unclassified")) %>%
  mutate(Percentage = round(as.numeric(Freq) / sum(Freq), 2 ) * 100)%>%
  knitr::kable(col.names = c("Broad Category", "Count", "Percentage"),
               caption = "Nonprofits in Comparison Set by Broad Category",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")
```

\newpage

```{r NTEE, echo=FALSE, warning=F, message=F}
tab.temp.ntee <- tab.ntee %>% as.data.frame()

colnames(tab.temp.ntee) <- c("NTEE", "Freq")

tab.temp.ntee %>%
  as.data.frame() %>%
  dplyr::mutate(NTEE= dplyr::recode(NTEE,
          "A" = "Arts, Culture, and Humanities" , 
          "B" =  "Education", 
          "C" = "Environment", 
          "D" = "Animals",
          "E" =  "Health Care",
          "F" = "Mental Health & Crisis Intervention" ,
          "G" = "Voluntary Health Associations & Medical Disciplines",
          "H" = "Medical Research",
          "I" = "Crime & Legal-Related" , 
          "J" = "Employment",
          "K" = "Food, Agriculture & Nutrition", 
          "L" = "Housing & Shelter", 
          "M" = "Public Safety, Disaster Preparedness & Relief" , 
          "N" = "Recreation & Sports", 
          "O" = "Youth Development", 
          "P" =  "Human Services", 
          "Q" = "International, Foreign Affairs & National Security" , 
          "R" = "Civil Rights, Social Action & Advocacy", 
          "S" = "Community Improvement & Capacity Building", 
          "T" = "Philanthropy, Voluntarism & Grantmaking Foundations", 
          "U" = "Science & Technology" , 
          "V" = "Social Science", 
          "W" = "Public & Societal Benefit", 
          "X" = "Religion Related", 
          "Y" = "Mutual & Membership Benefit" , 
          "Z" = "Unknown, Unclassified" )) %>%
  mutate(Percentage = round(as.numeric(Freq) / sum(Freq), 2 ) * 100) %>%
  arrange(NTEE) %>%
  knitr::kable(col.names = c("Major Group", "Count", "Percentage"),
               caption = "Nonprofits in Comparison Set by Major Group",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")

```

\newpage 

```{r LocationType, echo=FALSE, warning=F, message=F}
dat.filtered %>%
  select(LocationType) %>%
  group_by(LocationType) %>%
  summarize(Count = n()) %>%
  mutate(Percent = round(Count / dim(dat.filtered)[1], 2)* 100) %>%
  knitr::kable(caption = "Nonprofits in Comparison Set by City Type",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")

```



```{r Gender, echo=FALSE, warning=F, message=F}
dat.filtered %>%
  select(Gender) %>%
  group_by(Gender) %>%
  summarize(Count = n()) %>%
  mutate(Percent = round(Count / dim(dat.filtered)[1], 2)* 100) %>%
  knitr::kable(caption = "Nonprofits in Comparison Set by Gender",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")

```


```{r States, echo=FALSE, warning=F, message=F}
tab.state <- dat.filtered %>%
  select(State) %>%
  group_by(State) %>%
  summarize(Count = n()) %>%
  mutate(Percent = round(Count / dim(dat.filtered)[1], 2)* 100)

list(tab.state[1:25, ], tab.state[26:50, ]) %>%
  knitr::kable(caption = "Nonprofits in Comparison Set by State",
               #booktabs = T,
               format = "latex") %>%
  kableExtra::kable_styling(bootstrap_options = "striped",
                            full_width = F,
                            latex_options = "hold_position")

```






\newpage
\newpage

```{r CEOhist, echo=FALSE, warning=F, message=F}

ggplot(data = dat.filtered) +
  geom_histogram(aes(x = CEOCompensation), 
                 binwidth = 50000,
                 fill = "#1696d2") +
  ggtitle("Histogram of CEO Pay in Your Comparison Set") +
  ggplot2::xlab("CEO Total Pay") + 
  ggplot2::ylab("Count") +
  ggplot2::xlim(0, 1000000) + 
  labs(caption = paste("Note, there were", sum(dat.filtered$CEOCompensation > 1000000) , "nonprofits who paid there CEO more than $1,000,000 that are not displayed on this plot."))
  
```

\newpage 

```{r CorrPlot, echo=F, warning=F, message=F}
# plot(dat.filtered, 
#      main = "Correlation of Variables")

```

\newpage
```{r RelativePlots, echo=F, warning=F, message=F }

wanted.cols <- c("TotalExpense", "TotalEmployee", "TotalAssests", "GrossReceipts", "CEOCompensation")
vertical.bars <- data.frame( Vars =  wanted.cols,
                             Vals = unname(unlist((test.org[ , wanted.cols]))))


estimated.pay <- data.frame( Vars = "CEOCompensation" ,
                             Vals = weighted.average)


#Where is Your Org compared to other orgs? 
results %>%
  select(TotalExpense, TotalEmployee, TotalAssests, GrossReceipts, EIN, CEOCompensation, dist) %>%
  filter(TotalExpense <= quantile(TotalExpense, 0.95)) %>%
  #filter(TotalEmployee <= quantile(TotalEmployee, 0.95)) %>%
  filter(TotalAssests <= quantile(TotalAssests, 0.95)) %>%
  filter(GrossReceipts <= quantile(GrossReceipts, 0.95)) %>%
  filter(CEOCompensation <= quantile(CEOCompensation, 0.95)) %>%
  pivot_longer(!"EIN", names_to = "Vars", values_to = "Vals") %>%
  ggplot(aes(x = Vals)) + 
  geom_histogram()+
  geom_vline(aes(xintercept = Vals), data = vertical.bars, col = "red") +
  geom_vline(aes(xintercept = Vals), data = estimated.pay, col = "blue") +
  facet_wrap( ~Vars, scales = "free", ncol =2) +
  ggtitle("Where is your Org compared to the Orgs in the Comparison Set") +
  labs(subtitle = "Lower 95% of each variable is displayed",
       caption = "red = actual org values, blue = weighted average")


```

```{r CEOPlot, echo=F, warning=F, message=F }
results %>%
  select(CEOCompensation) %>%
  filter(CEOCompensation <= quantile(CEOCompensation, 0.95)) %>%
  ggplot(aes(x = CEOCompensation, y = ..density..)) + 
  geom_density() + 
  geom_vline(xintercept = test.org$CEOCompensation, col = "red") +
  geom_vline(xintercept = weighted.average, col = "blue") +
  geom_vline(xintercept = suggested.range[1], col = "green") +
  geom_vline(xintercept = suggested.range[2], col = "green") +
  labs(title = "Actual vs. Suggested CEO Pay",
              subtitle = "Lower 95% of data displayed",
              caption = "red = actual pay, blue = weighted average, green = suggested pay range")+
  xlab("CEO Pay") +
  ylab(" ")


```

\newpage

```{r DistanceMatrix, echo=F, warning=F, message=F }
#Average Distance score by attribute 
D.average <- colSums(distance.matrix) / dim(distance.matrix)[1]


#Average percent of total score by attribute 
dists <- results$dist

distance.matrix %>%
  pivot_longer(everything(), names_to = "Attribute", values_to = "Value") %>%
  ggplot(aes(x = Value)) +
  geom_histogram(binwidth = 0.05) +
  facet_wrap( ~ Attribute, scales = "free") +
  ggtitle("Histogram of Distance of Each Attribute")


dist.data.plot <- 
  distance.matrix %>%
  mutate(total = results$dist) %>%
  pivot_longer(!total, names_to = "attribute", values_to = "partial.distance") %>%
  mutate(prop = partial.distance/total) 

dist.data.plot %>%
  ggplot(aes(x = prop, y = total)) +
  geom_point(alpha = 0.1, size = 1)+
  facet_wrap(~ attribute) + 
  ggtitle("(Partial Distance / Total Distance) vs. Total Distance ")


dist.data.plot %>%
  ggplot(aes(x = partial.distance, y = total)) +
  geom_point(alpha = 0.1, size = 1)+
  facet_wrap(~ attribute) + 
  ggtitle("Partial Distance vs. Total Distance")


```

